Research Report

### 2023-02-28 tex/paper-acts/main.tex
Progress updating paper

@@

@@

ATLAS analyses use \abseta~2.5. After the HL-LHC upgrade, ATLAS will use \abseta~4. The goal of this research is to improve the efficiency of tracking and reconstruction, both in accuracy of reconstruction and computational time efficiency.
\subsection{pp at high $|\eta|$}

@@

\figtab{l l}{Truth Seed Range $\pt$ [MeV] & 500\\
r [mm] & 200\\
\Delta r & Optimized}

@@

\subsection{Kalman Filter algorithm}

@@

\subsection{Track fitting}

@@

\begin{enumerate}
\item Pileup [\#]: The number of soft processes to simulate alongside one hard process.
\item CMS Energy [GeV]: The center of mass energy of the symmetric pp collision.
\item QCD parameters: The QCD parameters to put into the event generator. Always ``Top:qqbar2ttbar=on''.
\item Gaussian Vertex $\sigma$ [$\sigma_t$, $\sigma_x$, $\sigma_y$, $\sigma_z$]: The standard deviation of the gaussian vertex from the center of the coordinate system (x,y,z) = (0,0,0). Always (5 ns, 0.0125 mm, 0.0125 mm, 55.5 mm).
\item B Field $B_x$,$B_y$,$B_z$ [T,T,T]: The magnetic field. Always 0, 0, 2.
\item Events per sequence [n]: The number of events to simulate per CKF reconstruction algorithm. Always 1.
\item FATRAS particle selection $\pt$ [GeV]: The minimum $\pt$ for FATRAS to propagate a particle. Always 150 MeV.
\item Truth seed $\pt$ [GeV]: The minimum $\pt$ for the truth seed algorithm. Always 1 GeV.
\item Truth seed number of hits [\#]: The minimum number of hits for the seed finder to declare a track.
\item Particle smearing pRel: Always 0.01. The relative momentum resolution used to smear the particle truth information to create track states.
\item Seed finder r: 200 mm. The bin side of the space point grid.
\item Seed finder $\Delta r$ min, max: 29, 52 mm. The min (max) distance in r between two measurements within one seed.
\item Seed finder collision region: -250 to 250 mm. The limiting location of collision region in z.
\item Seed finder z: -2000 mm to 2000 mm. The limiting location of measurements.
\item Seed finder max seeds per middle space point: 0. The number of middle space points that can be assigned per space point.
\item Seed finder max cot theta: 5.16. Maximum cot theta angle.
\item Seed finder scattering $\sigma$: 46.312. The maximum number of sigmas of scattering angle that should be searched.
\item Seed finder pT range: 500 MeV to 25.4 GeV. The min (max) pt range of scattering.
\item Seed finder b Field in Z direction: 1.99724 T. B field in z direction.
\item Seed finder impact radius max: 0.19345. The max impact parameter to search.
\item Track parameter estimation delta R: 10 mm, 100 mm. The min (max) deltaR between space points in a seed.
\item CKF minimum $\pt$: 1 GeV. The minimum pT of a track required in order to consider the track for efficiency measurements.
\item CKF minimum number of measurements: 6. The minimum number of hits in a track required in order to consider the track for efficiency measurements.
\end{enumerate}

@@

@@

Max scattering $p_{\textrm{T}}$ [GeV] & 10.0\\

@@

}{Efficiencies calculated from N ACTS sequence, 100 events per sequence, 100 pileup per event.}
\figpic{9cm}{4.png}{Efficiency plots as function of $\eta$, $\phi$, and $\pt$}
\subsection{$\eta$ from -3 to 3}

@@

\e{Score = particle efficiency - cur_fakerate_tracks - cur_duplicaterate_tracks / 5 - cur_runtime / 5}

@@

100 trials. The score function is

@@

@@

For the best trial,
\figpic{7cm}{/home/user1/time/2023-02-22/2023-02-22-13-30/compare-4/dup.png}{Plot of score per trial}
\figpic{7cm}{/home/user1/time/2023-02-22/2023-02-22-13-30/compare-4/eff.png}{Parameters per trial.}
\figpic{7cm}{/home/user1/time/2023-02-22/2023-02-22-13-30/compare-4/fak.png}{Efficiency per trial.}

@@

@@

For the best trial,
\figpic{7cm}{/home/user1/time/2023-02-23/2023-02-23-04-19/compare-4/dup.png}{Plot of score per trial}
\figpic{7cm}{/home/user1/time/2023-02-23/2023-02-23-04-19/compare-4/eff.png}{Parameters per trial.}
\figpic{7cm}{/home/user1/time/2023-02-23/2023-02-23-04-19/compare-4/fak.png}{Efficiency per trial.}

@@

@@

For the best trial,
\figpic{7cm}{/home/user1/time/2023-02-23/2023-02-23-02-45/compare-2/dup.png}{Track duplicate fraction as function of $\eta$, $\phi$, and $\pt$.}
\figpic{7cm}{/home/user1/time/2023-02-23/2023-02-23-02-45/compare-2/eff.png}{Track efficiency fraction as function of $\eta$, $\phi$, and $\pt$.}
\figpic{7cm}{/home/user1/time/2023-02-23/2023-02-23-02-45/compare-2/fak.png}{Track fake fraction as function of $\eta$, $\phi$, and $\pt$.}

@@

For the best trial,
\figpic{7cm}{/home/user1/time/2023-02-23/2023-02-23-10-07/compare-1/dup.png}{Track duplicate fraction as function of $\eta$, $\phi$, and $\pt$.}
\figpic{7cm}{/home/user1/time/2023-02-23/2023-02-23-10-07/compare-1/eff.png}{Track efficiency fraction as function of $\eta$, $\phi$, and $\pt$.}
\figpic{7cm}{/home/user1/time/2023-02-23/2023-02-23-10-07/compare-1/fak.png}{Track fake fraction as function of $\eta$, $\phi$, and $\pt$.}

@@

For the best trial,
\figpic{7cm}{/home/user1/time/2023-02-24/2023-02-24-07-07/compare-2/dup.png}{Track duplicate fraction as function of $\eta$, $\phi$, and $\pt$.}
\figpic{7cm}{/home/user1/time/2023-02-24/2023-02-24-07-07/compare-2/eff.png}{Track efficiency fraction as function of $\eta$, $\phi$, and $\pt$.}
\figpic{7cm}{/home/user1/time/2023-02-24/2023-02-24-07-07/compare-2/fak.png}{Track fake fraction as function of $\eta$, $\phi$, and $\pt$.}

>>>>

\geometry{top=0.75in, bottom=1in, left=0.75in, right=0.75in}

@@

\n{runtime}

@@

ATLAS analyses use \abseta~2.5. After the HL-LHC upgrade, analyses may use \abseta~4. The goal of this research is to improve the efficiency of tracking and reconstruction, both in accuracy of reconstruction and computational time efficiency.
% \subsection{pp at high $|\eta|$}

@@

\figtab{l l}{Truth Seed Range $\pt$ [MeV] & 500\\
r [mm] & 200\\
$\Delta$ r & Optimized}

@@

% \subsection{Kalman Filter algorithm}

@@

% \subsection{Track fitting}

@@

\begin{enumerate}
\item Pileup [\#]: The number of soft processes to simulate alongside one hard process.
\item CMS Energy [GeV]: The center of mass energy of the symmetric pp collision.
\item QCD parameters: The QCD parameters to put into the event generator. Always ``Top:qqbar2ttbar=on''.
\item Gaussian Vertex $\sigma$ [$\sigma_t$, $\sigma_x$, $\sigma_y$, $\sigma_z$]: The standard deviation of the gaussian vertex from the center of the coordinate system (x,y,z) = (0,0,0). Always (5 ns, 0.0125 mm, 0.0125 mm, 55.5 mm).
\item B Field $B_x$,$B_y$,$B_z$ [T,T,T]: The magnetic field. Always 0, 0, 2.
\item Events per sequence [n]: The number of events to simulate per CKF reconstruction algorithm. Always 1.
\item FATRAS particle selection $\pt$ [GeV]: The minimum $\pt$ for FATRAS to propagate a particle. Always 150 MeV.
\item Truth seed $\pt$ [GeV]: The minimum $\pt$ for the truth seed algorithm. Always 1 GeV.
\item Truth seed number of hits [\#]: The minimum number of hits for the seed finder to declare a track.
\item Particle smearing pRel: Always 0.01. The relative momentum resolution used to smear the particle truth information to create track states.
\item Seed finder r: 200 mm. The bin side of the space point grid.
\item Seed finder $\Delta r$ min, max: 29, 52 mm. The min (max) distance in r between two measurements within one seed.
\item Seed finder collision region: -250 to 250 mm. The limiting location of collision region in z.
\item Seed finder z: -2000 mm to 2000 mm. The limiting location of measurements.
\item Seed finder max seeds per middle space point: 0. The number of middle space points that can be assigned per space point.
\item Seed finder max cot theta: 5.16. Maximum cot theta angle.
\item Seed finder scattering $\sigma$: 46.312. The maximum number of sigmas of scattering angle that should be searched.
\item Seed finder pT range: 500 MeV to 25.4 GeV. The min (max) pt range of scattering.
\item Seed finder b Field in Z direction: 1.99724 T. B field in z direction.
\item Seed finder impact radius max: 0.19345. The max impact parameter to search.
\item Track parameter estimation delta R: 10 mm, 100 mm. The min (max) deltaR between space points in a seed.
\item CKF minimum $\pt$: 1 GeV. The minimum pT of a track required in order to consider the track for efficiency measurements.
\item CKF minimum number of measurements: 6. The minimum number of hits in a track required in order to consider the track for efficiency measurements.
\end{enumerate}

@@

The procedure for optimization is
\begin{enumerate}
\item Initialize the optimizer and set the first trial paremters to \fref{192837a}.
\item Run N trials using the following steps.
\begin{enumerate}
\item Run one ACTS sequence with one event generated in the ODD.:
\begin{enumerate}
\item Pythia8, with multiple pileup particles.
\item GEANT+Fatras with the ODD.
\item Digitization with ACTS.
\item Seeding with ACTS.
\item CKF track reconstruction wth ACTS.
\end{enumerate}
\item Calculate the score, calculate new parameters.
\end{enumerate}
\item Report the N trials, their parameters, and the score of each trial.
\end{enumerate}

@@

Max scattering $\pt$ [GeV] & 10.0\\

@@

}{Initial efficiencies calculated from 1 ACTS sequence, with 100 events, with 100 pileup particles per event.}
\figpic{4cm}{/home/user1/time/2023-02-23/2023-02-23-11-14/single/dup.png}{Initial track duplicate fraction as function of $\eta$, $\phi$, and $\pt$.}
\figpic{4cm}{/home/user1/time/2023-02-23/2023-02-23-11-14/single/eff.png}{Initial track efficiency fraction as function of $\eta$, $\phi$, and $\pt$.}
\figpic{4cm}{/home/user1/time/2023-02-23/2023-02-23-11-14/single/fak.png}{Initial track fake fraction as function of $\eta$, $\phi$, and $\pt$.}

@@

\e{\sco = \preff - \tfak - \tdup / 5 - \runtime / 5}

@@

@@

Max \textrm{cot}($\theta$) [\textrm{cot}($\theta$)] & 9.97960491147606\\
Max $\Delta R$ [mm] & 193.7852042641601\\
Min $\Delta R$ [mm]  & 6.186148583682443\\
Max $r_\textrm{impact}$ [mm] & 5.3520710118467525\\
Max scattering $p_{\textrm{T}}$ [GeV] & 6.270284198944674\\
Max Seeds Per Middle Space Point [n] & 6\\
Radiation Length Per Seed & 0.0012607157984066096\\
Scattering $\sigma$ [$\sigma$] & 40.60867756209729

@@

The parameters from the trial with the highest score were used to generate 100 events, with 100 pileup particles each. The resulting efficiency plots are below.,
\figpic{4cm}{/home/user1/time/2023-02-22/2023-02-22-13-30/compare-4/dup.png}{Track duplicate fraction as function of $\eta$, $\phi$, and $\pt$.}
\figpic{4cm}{/home/user1/time/2023-02-22/2023-02-22-13-30/compare-4/eff.png}{Track efficiency fraction as function of $\eta$, $\phi$, and $\pt$.}
\figpic{4cm}{/home/user1/time/2023-02-22/2023-02-22-13-30/compare-4/fak.png}{Track fake fraction as function of $\eta$, $\phi$, and $\pt$.}

@@

Max \textrm{cot}($\theta$) [\textrm{cot}($\theta$)] & 9.955015534804646\\
Max $\Delta R$ [mm] & 92.55025939304613\\
Min $\Delta R$ [mm] & 4.954090296295913\\
Max $r_\textrm{impact}$ [mm] & 6.220456981531933\\
Max scattering $p_{\textrm{T}}$ [GeV] & 25.742954905376422\\
Max Seeds Per Middle Space Point [n] & 9\\
Radiation Length Per Seed & 0.050182305379654776\\
Scattering $\sigma$ [$\sigma$] & 41.62774067749547 

@@

The parameters from the trial with the highest score were used to generate 100 events, with 100 pileup particles each. The resulting efficiency plots are below.,
\figpic{4cm}{/home/user1/time/2023-02-23/2023-02-23-04-19/compare-4/dup.png}{Track duplicate fraction as function of $\eta$, $\phi$, and $\pt$.}
\figpic{4cm}{/home/user1/time/2023-02-23/2023-02-23-04-19/compare-4/eff.png}{Track efficiency fraction as function of $\eta$, $\phi$, and $\pt$.}
\figpic{4cm}{/home/user1/time/2023-02-23/2023-02-23-04-19/compare-4/fak.png}{Track fake fraction as function of $\eta$, $\phi$, and $\pt$.}

@@

Max \textrm{cot}($\theta$) [\textrm{cot}($\theta$)] & 8.585756197102755\\
Max $\Delta R$ [mm] & 226.44349889012312\\
Min $\Delta R$ [mm] & 13.667849506894473\\
Max $r_\textrm{impact}$ [mm] & 21.905395073424017\\
Max scattering $p_{\textrm{T}}$ [GeV] & 49.44440486374715\\
Max Seeds Per Middle Space Point [n] & 10\\
Radiation Length Per Seed & 0.029552672847034882\\
Scattering $\sigma$ [$\sigma$] & 17.750322166919286 

@@

The parameters from the trial with the highest score were used to generate 100 events, with 100 pileup particles each. The resulting efficiency plots are below.,
\figpic{4cm}{/home/user1/time/2023-02-23/2023-02-23-02-45/compare-2/dup.png}{Track duplicate fraction as function of $\eta$, $\phi$, and $\pt$.}
\figpic{4cm}{/home/user1/time/2023-02-23/2023-02-23-02-45/compare-2/eff.png}{Track efficiency fraction as function of $\eta$, $\phi$, and $\pt$.}
\figpic{4cm}{/home/user1/time/2023-02-23/2023-02-23-02-45/compare-2/fak.png}{Track fake fraction as function of $\eta$, $\phi$, and $\pt$.}

@@

The parameters from the trial with the highest score were used to generate 100 events, with 100 pileup particles each. The resulting efficiency plots are below.,
\figpic{4cm}{/home/user1/time/2023-02-23/2023-02-23-10-07/compare-1/dup.png}{Track duplicate fraction as function of $\eta$, $\phi$, and $\pt$.}
\figpic{4cm}{/home/user1/time/2023-02-23/2023-02-23-10-07/compare-1/eff.png}{Track efficiency fraction as function of $\eta$, $\phi$, and $\pt$.}
\figpic{4cm}{/home/user1/time/2023-02-23/2023-02-23-10-07/compare-1/fak.png}{Track fake fraction as function of $\eta$, $\phi$, and $\pt$.}

@@

The parameters from the trial with the highest score were used to generate 100 events, with 100 pileup particles each. The resulting efficiency plots are below.,
\figpic{4cm}{/home/user1/time/2023-02-24/2023-02-24-07-07/compare-2/dup.png}{Track duplicate fraction as function of $\eta$, $\phi$, and $\pt$.}
\figpic{4cm}{/home/user1/time/2023-02-24/2023-02-24-07-07/compare-2/eff.png}{Track efficiency fraction as function of $\eta$, $\phi$, and $\pt$.}
\figpic{4cm}{/home/user1/time/2023-02-24/2023-02-24-07-07/compare-2/fak.png}{Track fake fraction as function of $\eta$, $\phi$, and $\pt$.}
